---
title: "Viz of least and most compact lakes"
author: "Nicole Smith"
date: '2018-01-22'
output:
  html_document: default
  html_notebook: default
---
```{r setup, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE}
library(tidyverse)
library(sf)
library(leaflet)

# lakes_poly <- st_read("D:/Continental_Limnology/Data_Working/LAGOS_US_Predecessors.gdb", "NHDWaterbody_LAGOS", stringsAsFactors = FALSE) %>%
#   st_zm(drop=TRUE) %>%
#   filter(AreaHa >= 4) %>%
#   mutate(FCode = factor(FCode)) %>%
#   mutate(FType = factor(FType)) %>%
#   select(Permanent_Identifier, GNIS_Name, AreaSqKm, ReachCode, FType, FCode, STATE, lagoslakeid, Shape_Area, Shape_Length)
# 
# states <- st_read('C:/Users/smithn78/Dropbox/CL_HUB_GEO/QAQC/LAGOS_US_GIS_Data_v0.2.gdb','STATE') %>%
#   st_simplify(TRUE, 5000)
# 
# save.image("C:/Users/smithn78/Dropbox/ContLimno Nicole/MyR/multilake_plots/compactness.RData")

load('compactness.RData')
```

I'll use the `multiplot` function from the [Cookbook for R](http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/) site with an [adaptation to add a title](http://www.guru-gis.net/multiplot-function-for-ggplot/).
```{r functions, echo=FALSE, warning=FALSE, message=FALSE}
## From R cookbook
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL, title="") {
  require(grid)
 
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
 
  numPlots = length(plots)
 
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
 
  if (nchar(title)>0){
    layout<-rbind(rep(0, ncol(layout)), layout)
  }
 
  if (numPlots==1) {
    print(plots[[1]])
 
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout), heights =if(nchar(title)>0){unit(c(0.5, rep(5,nrow(layout)-1)), "null")}else{unit(c(rep(5, nrow(layout))), "null")} )))
 
    # Make each plot, in the correct location
    if (nchar(title)>0){
      grid.text(title, vp = viewport(layout.pos.row = 1, layout.pos.col = 1:ncol(layout)))
    }
 
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
 
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# With help from http://rstudio-pubs-static.s3.amazonaws.com/2852_379274d7c5734f979e106dcf019ec46c.html
stage_plots <- function(data, region=NULL, size = c("maxzoom", "scale")) {
  size = match.arg(size)
  !is.null(region) || stop("you must supply a region identifier")
  if (size == "scale") {
    boxes <- list()
    for (i in 1:nrow(data)) {
      boxes[[i]] <- st_bbox(data[i,])
    }
    x_max_range <- max(as.vector(sapply(boxes, function(x) x['xmax'] - x['xmin'])))
    y_max_range <- max(as.vector(sapply(boxes, function(x) x['ymax'] - x['ymin'])))
    max_dim <- max(y_max_range, x_max_range)
  }
  plots <- list()
  for (i in 1:nrow(data)) {
    plot_item <- data[i,] %>% as("Spatial") %>% fortify(region=region)
    plot <- ggplot(plot_item) + 
      geom_polygon(aes(x=long, y=lat, group=group), fill = 'lightsteelblue3', color = 'dodgerblue4') +
      theme_void() +
      coord_fixed()
    if (size == "scale") {
      centroid <- data[i,] %>% st_centroid() %>% st_coordinates()
      x_limits <- c(centroid[,'X'] + .5*max_dim, centroid[,'X'] - .5*max_dim)
      y_limits <- c(centroid[,'Y'] + .5*max_dim, centroid[,'Y'] - .5*max_dim)
      plot <- plot +
        xlim(x_limits) + 
        ylim(y_limits)
    }
    plots[[i]] <- plot
  }
  return(plots)
}

```

I'll be using the *isoperimetric quotient*, a common measure of compactness that compares the area of the shape to the area of a circle with the same perimeter. $$\frac{4\pi A}{P^2}$$With this measure, a value of 1 indicates the lake is nearly circular, and lower values represent less compactness.
```{r}
lakes <- lakes_poly %>%
  mutate(compactness = (4*pi*Shape_Area)/(Shape_Length**2)) %>%
  arrange(-compactness) %>%
  mutate(compact_rank = row_number())
```

Prepare a few subsets for plotting.
```{r}
most_compact_100 <- lakes %>%
  top_n(-100, compact_rank)

least_compact_100 <- lakes %>%
  top_n(100, compact_rank)

midpoint <- floor(nrow(lakes)/2)

mid_compact_100 <- lakes %>%
  filter(between(compact_rank, midpoint-50, midpoint+49))
```

#100 most compact lakes in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
staged1 <- stage_plots(most_compact_100, "lagoslakeid")
p1 <- multiplot(plotlist=staged1, cols=10)
```

#100 least compact lakes in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
staged2 <- stage_plots(least_compact_100, "lagoslakeid")
p2 <- multiplot(plotlist=staged2, cols=10)
```

#100 lakes of median compactness in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
staged3 <- stage_plots(mid_compact_100, "lagoslakeid")
p3 <- multiplot(plotlist=staged3, cols=10)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Prep the layers for leaflet
proj_string <- "+proj=longlat +datum=WGS84"
L.most_compact_100 <- most_compact_100 %>%
  st_transform(proj_string)
L.least_compact_100 <- least_compact_100 %>%
  st_simplify(dTolerance=200) %>%
  st_transform(proj_string)
L.states <- states %>%
  st_transform(proj_string)
```

#Where are the 100 **most** compact lakes?
```{r, echo=FALSE, warning=FALSE, out.width="100%"}
popup_template <- "<b>GNIS Name:</b><br/>%s<br/><b>FCode:</b><br/>%s"

(m1 <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery, options=providerTileOptions(opacity=.75)) %>%
  addPolygons(data=L.states, weight = 2, opacity = .25, color="white", fill=NULL) %>%
  addPolygons(data=L.most_compact_100, weight = 2) %>%
  addCircleMarkers(data = st_centroid(L.most_compact_100), radius = 4, fillOpacity = 1, stroke = FALSE,
                    popup = ~sprintf(popup_template, GNIS_Name, FCode))
   )
```

#Where are the 100 **least** compact lakes?
```{r, echo=FALSE, warning=FALSE, out.width="100%"}
(m2 <- leaflet() %>%
   addProviderTiles(providers$Esri.WorldImagery, options=providerTileOptions(opacity=.75)) %>%
   addPolygons(data=L.states, weight = 2, opacity=.25, color="white", fill=NULL) %>%
   addPolygons(data=L.least_compact_100, weight = 2) %>%
   addCircleMarkers(data = st_centroid(L.least_compact_100), radius = 4, fillOpacity = 1, stroke = FALSE,
                    popup = ~sprintf(popup_template, GNIS_Name, FCode))
   )

```


#Are there compactness patterns across the United States?
```{r, echo=FALSE, warning=FALSE, out.width="100%"}
lake_points <- st_centroid(lakes) %>%
  mutate(x = st_coordinates(.)[,"X"]) %>%
  mutate(y = st_coordinates(.)[,"Y"])

ggplot() +
  stat_summary_hex(data = lake_points, aes(x, y,  z=compactness), binwidth = c(50000, 50000)) +
  geom_sf(data = states, color = 'grey50', fill = NA, size = .5) +
  scale_fill_gradient2(midpoint=median(lake_points$compactness), name = "compactness") +
  theme_void()

```

#Distribution of compactness.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
(hist <- ggplot(data=lakes, aes(x=compactness)) +
   geom_histogram(alpha = .5, fill = "blue") +
   geom_vline(aes(xintercept=median(compactness)))
)
```


